Environmental Context Dependency in Species Interactions

### 2018-08-21 EDM_SNI_analysis.Rmd
molly's icons and manuscript abstract

```{r setup, include=FALSE,message=F}
# Load required packages
require(tidyverse)
require(rEDM)
require(quantreg)
require(knitr)
require(kableExtra)
require(extrafont)
require(GGally)
require(ggalt)
require(sna)
require(network)
require(gridExtra)
require(RANN)

@@

```{r interactions density plot by type,fig.height=6,fig.width=6}
# Density plot
interaction_distribution_by_type <- spp.only.mod.coeffs %>%
  filter(!is.na(type),type != "intraspecies")%>%
  group_by(type) %>%
  mutate(mean_int=mean(value,na.rm=T))%>%
  ungroup()%>%
  ggplot(aes(value,..density..))+
  geom_density(aes(fill=type),alpha=0.8,col=NA)+
  geom_vline(xintercept=0,linetype=2)+
  geom_vline(aes(xintercept=mean_int))+
  scale_fill_manual(values=c("darkorange1","orangered2","darkorchid1","palegreen3","skyblue1"),name="")+
   guides(fill=F)+
  xlim(-1.5,1.5)+
  labs(x="Interaction strength",y="Probability density")+
  geom_text(aes(x=-1.5,y=0.5,label=type),hjust="inward",family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
  facet_grid(rows=vars(type))+
  theme(strip.text = element_blank())

>>>>

```{r setup, include=FALSE,message=F}
# Load required packages
require(tidyverse)
require(here)
require(rEDM)
require(quantreg)
require(knitr)
require(kableExtra)
require(extrafont)
require(png)
require(ggridges)
require(GGally)
require(ggalt)
require(ggsci)
require(sna)
require(network)
require(gridExtra)
require(RANN)

@@

```{r interactions density plot by type,fig.height=6,fig.width=6}
# Density plot
require(png)
a1 <- readPNG(here::here("plots","icons","macrocystis-transparent.png")) %>% rasterGrob()
interaction_distribution_by_type <- spp.only.mod.coeffs %>%
  filter(!is.na(type),type != "intraspecies")%>%
  group_by(type) %>%
  mutate(mean_int=mean(value,na.rm=T))%>%
  ungroup()%>%
  ggplot(aes(value,type,fill=type))+
  geom_density_ridges(alpha=0.8,scale=1.4)+
  # stat_density_ridges(geom="density_ridges_gradient",calc_ecdf='true',quantiles=4,alpha=0.8,scale=1)+
  # scale_fill_viridis_d(name="quantiles",alpha=0.8)+
  geom_vline(xintercept=0,linetype=2)+
  # geom_vline(aes(xintercept=mean_int,color=type))+
  scale_fill_manual(values=c("darkorange1","orangered2","darkorchid1","palegreen3","skyblue1"),name="")+
  scale_color_manual(values=c("darkorange1","orangered2","darkorchid1","palegreen3","skyblue1"),name="")+
  guides(fill=F,color=F)+
  xlim(-1.5,1.5)+
  labs(x="Interaction strength",y="Probability density")+
  annotation_custom(a1,ymin=2,ymax=3,xmin=-1,xmax=-0.5)+
  # geom_text(aes(x=-1.5,y=0.5,label=type),hjust="inward",family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
  # facet_grid(rows=vars(type))+
  theme(strip.text = element_blank())

### 2018-08-21 manuscript.Rmd
molly's icons and manuscript abstract

```{r packages and workspace}
# # header-includes:
#   - \usepackage[left]{lineno}
#   - \linenumbers
#   - \renewcommand\linenumberfont{\normalfont\bfseries\small}
require(tidyverse)
require(extrafont)
load("ws/wrkspce_07.20.18.RData")

@@

Ecological interactions are not uniform across time, and instead vary with environmental conditions. Interactions among species are often measured with short-term controlled experiments, but these experiments are subject to the particular environmental conditions under which they are performed. As an alternative, we utilize empirical dynamic modeling applied to a 30-year time series to estimate species interactions across a wide range of environmental conditions. We show that environmental context influences the strength and direction of species interactions. In so doing, we are able to confirm and extend results from previous studies, as well as identify potentially important but understudied dynamics, including the importance of specifically studying variation rather than mean interaction outcomes. The significant context dependency in species interactions found in this study argues for a greater utilization of long-term data and empirical dynamic modeling in studies of ecosystem dynamics.

>>>>

```{r packages and workspace}
# # header-includes:
#   - \usepackage[left]{lineno}
#   - \linenumbers
#   - \renewcommand\linenumberfont{\normalfont\bfseries\small}
require(tidyverse)
require(extrafont)
require(ggridges)
load("ws/wrkspce_07.20.18.RData")

@@

Ecological interactions are not uniform across time. They commonly vary with environmental conditions. Yet, interactions among species are often measured with short-term controlled experiments whose outcome can depend greatly on the particular environmental conditions under which they are performed. Running experiments repeatedly to address this context challenge is an inefficient solution at best. As an alternative, we utilize empirical dynamic modeling applied to a 30-year time series from coastal kelp forests to estimate species interactions across a wide range of environmental conditions. We show that environmental context greatly alters the strength and direction of species interactions. In so doing, we confirm and extend results from previous studies on the power of empirical dynamic modeling. We identify potentially important but understudied kelp forest dynamics, including the importance of specifically studying variation in interaction strength rather than mean interaction outcomes. The significant context dependency in species interactions found in this study argues for a greater utilization of long-term data and empirical dynamic modeling in studies of the dynamics of other ecosystems.
